---
title: "Summary Table"
author: "Jamie Ash"
date: "05/31/2022"
output: 
  bookdown::html_document2:  
    self-contained: yes
    theme: paper #cerulean #cosmo #journal #readable
    toc: false
    smooth_scroll: true
    toc_depth: 3
    toc_float: true
    number_sections: false
    fig_caption: yes
    code_folding: hide
---  

```{r setup, include=FALSE, message = FALSE}
source("libraries.R")
source("functions.R")
source("blooms.R")
rasterOptions(maxmemory = 120e+10, memfrac = 0.9)
knitr::opts_chunk$set(echo = TRUE)
```

# Data processing

Functions
```{r, message = FALSE}

# input is a raster output is a boolian raster of bloom/not bloom
bool = function(x){
  u <- calc(x, fun = median, na.rm = TRUE)
  o <- calc(x, fun = mad, na.rm = TRUE)
  boo <- x > (u + o)
  extent(boo) <- extent(x)
  boo <- setZ(boo, z = getZ(x), name = "time")
  boo
  }

# Input is 3D chla raster output is time series vector of bloom area
barea = function(b){
  t   = getZ(b)
  #boo = bool(x)
  rr  = reclassify(b, cbind(0,NA))
  a   = raster::area(rr)
  #aa  = a * boo
  temp <- raster::mask(a, rr)
  km2 = cellStats(temp, stat = "sum", na.rm = TRUE)
  data.frame(area = as.numeric(km2), time= t)
}

# input x is a raster of raw chl
# inpur x is a raster of 0/1 bloom non-bloom
bmag <- function(x, b){
  t = getZ(x)
  # create a raster of bloom 0, 1's as b
  # boo = bool(x)
  # temp = x * boo
  
  # Area time series for duration 
  rr <- reclassify(b, cbind(0,NA))
  # find area of each cell + mask cells that matter
  temp <- raster::mask(x, rr)
  
  mag  = cellStats(temp, stat ="max", na.rm = TRUE)
  data.frame(mag = as.numeric(mag), time= t)
  }  

# # input x is a raster of chl
# bcent <- function(x, b){
#   # create a raster of bloom 0, 1's as b
#   pix  <- cellStats(b, stat = sum, na.rm = TRUE)
#   ind  <- as.numeric(which(pix==max(pix, na.rm = TRUE)))
#   
#   # chla layer with largest bloom area
#   temp <- raster::mask(subset(x, ind), subset(b, ind))
#   
#   # index of cell with largest chla value. maybe I should do mean lat/lon 
#   # turn long format data frame. Remove NA values (not blooms). 
#   # Take mean of lon/lats
#   idx  <- which.max(temp)
#   s    <- dim(temp)
#   e    <- extent(temp)
#   lats <- seq(from = e[3], to = e[4], length = s[2])
#   lons <- seq(from = e[1], to = e[2], length = s[1])
#   grid <- meshgrid(lons, lats)
#   c(grid$X[idx[1]], grid$Y[idx[1]])
# }

vectorize  <- function(x) {
  sdate <- getZ(x)
  # x = raster::flip(x, direction = "x")
  # x = raster::flip(x, direction = "y")
  # 
  # e = extent(x)
  # x = t(x)
  # extent(x) = e
  # 
  x     <- rasterToPoints(x)
  x     <- data.frame(x)
  colnames(x) <- c("lon", "lat", as.character(sdate))
  x     <- reshape2::melt(x, id.vars = c("lat", "lon"))
  colnames(x) <- c("lats", "lons", "time", "val")
  x
}

fronts = function(ras,
                  downsize = 10,
                  I = 2){
  # ras  = raster::flip(ras, direction = "x")
  thresh = median(ras, na.rm = TRUE) + mad(ras, na.rm = TRUE)*I
  idx  = which(values(ras) > thresh)
  e    = extent(ras)
  s    = dim(ras)
  lons = seq(from = e[1], to = e[2], length = s[1]) 
  lats = seq(from = e[3], to = e[4], length = s[2]) 
  grid = pracma::meshgrid(lons, lats)
  fronts = data.frame(lons = grid$X[idx], lats = grid$Y[idx], value = ras[idx])
  idx = seq(1, nrow(fronts), downsize)
  fronts = fronts[idx, ]
  fronts
}

bcent <- function(ras){
  # create a raster of bloom 0, 1's as b
  pix  <- cellStats(ras, stat = mean, na.rm = TRUE)
  idx  <- as.numeric(which(pix == max(pix, na.rm = TRUE)))
  # temp = raster::mask(subset(ras, ind), subset(b, ind))
  
  # chla layer with largest bloom area
  temp = subset(ras, idx)
  # smooth that layer to reduce costal effects
  temp = raster.gaussian.smooth(x = temp, n = 21, sigma = 5)
  #temp = oreant(temp, t1 = TRUE)
  
  temp = vectorize(temp)
  idx = which(temp$val == max(temp$val, na.rm = TRUE))
  idx = max(idx)
  temp[idx,]
}

# bcent <- function(ras){
#   # create a raster of bloom 0, 1's as b
#   pix  <- cellStats(ras, stat = mean, na.rm = TRUE)
#   idx  <- as.numeric(which(pix == max(pix, na.rm = TRUE)))
#   # temp = raster::mask(subset(ras, ind), subset(b, ind))
#   # chla layer with largest bloom area
#   temp = subset(ras, idx)
# 
#   temp = fronts(temp, I = 3)
#   #temp$value = scale01(temp$value)
#   y = mean(temp$lats, na.rm = TRUE)
#   x = mean(temp$lons, na.rm = TRUE)
#   data.frame(lons = x, lats = y)
#   }
#   idx = which(temp$value == max(temp$value, na.rm = TRUE))
#   idx = max(idx)
#   temp[idx,]
# }


mymax = function(x, par = "area"){
    idx = which(x[, par] == max(x[, par], na.rm = TRUE))
    x = x[idx, ]
    x[1, par]
}

# connects blooms separated by desired duration
fillblips = function(x, dur = 2){
  blooms = rle(x)
  # index runs sorter than the desired duration
  shortblooms = blooms$lengths <= dur
  # index short that are false false runs
  deadair = shortblooms & !blooms$values
  # replace shorts with true
  blooms$values[deadair] = TRUE
  rep(blooms$values, blooms$lengths)
}

# removes all blooms shorter than the desired duration
# determines if element is a member of consecutive TRUE/FALSE
longblooms = function(x, dur = 2) {
  blooms    = rle(x)
  bigblooms = blooms$lengths > dur
  #bigblooms = blooms$values == TRUE
  bigblooms = rep(bigblooms, blooms$lengths)
  # TRUE * FALSE = FALSE so this filters the non blooms
  bigblooms = bigblooms & x
  #bigblooms = bigblooms == 1
  bigblooms
}

# input is a vector of true/false bloom/not bloom values
# output is a data frame of start and end indices of bloom times
bloomtails = function(x){
  # Compute endpoints of run
  blooms = rle(x)
  # subsetting only trues
  #truelengths = blooms$lengths[blooms$values]
  
  # subset trues after the fact
  end = cumsum(blooms$lengths)
  start = c(1, lag(end)[-1] + 1)
  
  data.frame(start, end)[blooms$values, ]
  } 

# x is a data frame of bloom start and end indexes from bloom tails function
# ras is a raster brick with a z as the time variable
idxbloom = function(x, ras){
  subset(ras, x[1]:x[2])
}

idxbloom = function(x, ras){
  t = getZ(ras)
  y = subset(ras, x[1]:x[2])
  t = t[x[1]:x[2]]
  setZ(y, z = t, name = "time")
  }

```

Higher level functions
```{r, message = FALSE, warning = FALSE}
filt_stl = function(ras = chl, res = 8, sub.start = 2){
  #res = 8 # 8 day resolution
  # needs at least 2 years to run
  ts = cellStats(ras, stat = "mean", na.rm = TRUE)
  ts = as.numeric(ts)
  
  so = stlplus(x   = ts,      # One time series of the raster
               t   = getZ(ras),       # datetime vector
               n.p = floor(365/res),      # give the period of seasonality
               s.window  = floor(30/res), # length of window: about a month 
               sub.start = sub.start, # if data does not start on XXXX-01-01
               outer     = 4) # idk something with the iterations
  
  raw  = so$data$raw
  anom = so$data$remainder
  clim = so$data$trend
  seas = so$data$seasonal
  seas_clim =  seas + clim
  
  data.frame(raw = raw, anom = anom, clim = clim, seas = seas, seas_clim =  seas + clim)
}

bloomspan = function(anom, time = getZ(chl), gaps = 0, dur = 1){
  # apply loess to 
  
  u   = median(anom, na.rm = TRUE)
  o   = mad(anom, na.rm = TRUE)
  boo = anom > u + o/2
  boo[is.na(boo)] = FALSE
  
  # Instead of filling blips, apply loess
  #boo = fillblips(boo, dur = 4)
  boo = fillblips(boo, dur = gaps)
  boo = longblooms(boo, dur = dur)
  
  bloom_idx = bloomtails(boo)
  
  dablooms  = data.frame(sdate = time[bloom_idx$start], 
                         edate = time[bloom_idx$end], 
                         duration = abs(time[bloom_idx$start] - time[bloom_idx$end]),
                         start = bloom_idx$start,
                         end = bloom_idx$end)
  idx_summer = month(dablooms$sdate) > 4 & month(dablooms$sdate) < 10 
  #bloom_idx  = bloom_idx[idx_summer, ]
  dablooms   = dablooms[idx_summer, ]
  dablooms
  }

sumstats = function(blooms = chlb, signal1 = chl, signal2 = chla, sdate, edate){
  
  ts = as.Date(sdate)
  te = as.Date(edate)
  tc = getZ(signal1)
  ta = getZ(signal2)
  
  nearest = function(x, y) which.min(abs(y - x))
  
  start = list()
  end   = list()
  for(i in 1:length(ts)) start[[i]] = as.Date(ts[i])
  for(i in 1:length(te)) end[[i]] = as.Date(te[i])
  
  start = lapply(X = start, FUN = nearest, y = tc)
  end   = lapply(X = end, FUN = nearest, y = tc) 
  
  idx = data.frame(start = unlist(start), end = unlist(end))
  #idx = tbl[,c("start", "end")]
  
  cb = apply(idx, MARGIN = 1, FUN = idxbloom, ras = blooms)
  
  # instead of calculating a 0/1 I can vectorize the blooms and turn them into 
  # long-format tables
  
  co = apply(idx, MARGIN = 1, FUN = idxbloom, ras = signal1)
  
  ca = apply(idx, MARGIN = 1, FUN = idxbloom, ras = signal2)
  
  # then apply max function to list of time series to get value
  #cents = mapply(bcent, ras = co, SIMPLIFY = FALSE)
  cents = mapply(bcent, ras = ca, SIMPLIFY = FALSE)
  cents = do.call(rbind, cents)
  
  mags  = mapply(bmag, x = co, b = cb, SIMPLIFY = FALSE)
  areas = lapply(cb,    FUN = barea)
  
  m     = lapply(mags,  FUN = mymax, par = "mag")
  a     = lapply(areas, FUN = mymax, par = "area")
  
  duration = as.Date(edate) - as.Date(sdate)
  
  sumtbl = data.frame(sdate    = as.Date(sdate),
                      edate    = as.Date(edate),
                      duration = duration,
                      mag      = round(unlist(m), 2), 
                      area     = round(unlist(a), 2),
                      lat      = cents[,1],
                      lon      = cents[,2])
  
  # idx = sumtbl$duration > 16
  # sumtbl = sumtbl[idx, ]
  sumtbl
}

# Fast furiuer series decomposiiton summing first n components
nff = function(x = NULL, n = NULL, up = 1L){
  #The direct transformation
  #The first frequency is DC, the rest are duplicated
  dff = fft(x)
  #The time
  t = seq(from = 1, to = length(x))
  #Upsampled time
  nt = seq(from = 1, to = length(x)+1-1/up, by = 1/up)
  #New spectrum
  ndff = array(data = 0, dim = c(length(nt), 1L))
  ndff[1] = dff[1] #Always, it's the DC component
  if(n != 0){
    ndff[2:(n+1)] = dff[2:(n+1)] #The positive frequencies always come first
    #The negative ones are trickier
    ndff[length(ndff):(length(ndff) - n + 1)] = dff[length(x):(length(x) - n + 1)]
  }
  #The inverses
  indff = fft(ndff/73, inverse = TRUE)
  idff = fft(dff/73, inverse = TRUE)
  ret = data.frame(time = nt, signal = Mod(indff))
  return(ret)
}

```

```{r}
# using a two degree box around aloha
lons = c(-158, -130)
lats = c(22, 35)
e = extent(lons, lats)
# ------------------------------------------------------------------------------
# Set variables
lat_varid = "lat"
lon_varid = "lon"
sdate = as.Date("1997-01-01")
edate = as.Date("2022-12-31")
var = "CHL"
url = "https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/cmems_obs-oc_glo_bgc-plankton_my_l4-multi-4km_P1M?"
origin = "1900-01-01"

data = nc_open(url, verbose = FALSE, write = FALSE)
lat  = ncvar_get(data, varid = lat_varid)
lon  = ncvar_get(data, varid = lon_varid)
time = ncvar_get(data, varid = "time")
time = as.Date(time, origin = origin)

idx_lat <- which(lat > lats[1] & lat < lats[2])
idx_lon <- which(lon > lons[1] & lon < lons[2])
idx_time <- which(time >= sdate & time <= edate)


idx_ras <- paste("CHL",
                 paste("[", range(idx_time)[1], ":1:", range(idx_time)[2], "]", sep = ""),
                 paste("[", range(idx_lat)[1],  ":1:", range(idx_lat)[2],  "]", sep = ""),
                 paste("[", range(idx_lon)[1],  ":1:", range(idx_lon)[2],  "]", sep = ""),
                 sep = "")

idx_time <- paste("time", paste("[", range(idx_time)[1], ":1:",range(idx_time)[2], "]", sep = ""), sep = "")
idx_lat  <- paste(lat_varid, paste("[", range(idx_lat)[1], ":1:",range(idx_lat)[2],  "]", sep = ""), sep = "")
idx_lon  <- paste(lon_varid, paste("[", range(idx_lon)[1], ":1:",range(idx_lon)[2],  "]", sep = ""), sep = "")
idx <- paste(idx_lat, idx_lon, idx_time, idx_ras, sep = ",")

url <- paste(url, idx, sep = "")

nc_close(data)
rm(data)

data = nc_open(url, verbose = FALSE, write = FALSE)

lat  = ncvar_get(data, varid = lat_varid)
lon  = ncvar_get(data, varid = lon_varid)
time = ncvar_get(data, varid = "time")
time = as.Date(time, origin = origin)
ras  = ncvar_get(data)
nc_close(data)

s = dim(ras)
ras = raster::brick(ras)
ras = t(ras)
ras = setZ(ras, z = as.Date(time, origin = org), name = "time")
extent(ras) = extent(min(lons), max(lons), min(lats), max(lats))

chl = ras
rm(ras)
gc()

```

Data download
```{r, eval = FALSE, message = FALSE, warning = FALSE}
blooms = blooms[2:nrow(blooms),]

lons = c(-170, -130)
lats = c( 17,   35)
 
# I had 2003-12-01 for the old sum table
sdate = as.Date("2003-01-02")
edate = as.Date("2021-12-01")
chl = loadchl(url = "https://upwell.pfeg.noaa.gov/erddap",
              id  = "erdMH1chla8day",
              lat = lats,
              lon = lons,
              sdate = sdate,
              edate = edate,
              large = FALSE)

chl = loadchl(url = "https://upwell.pfeg.noaa.gov/erddap",
              id  = "erdMH1chlamday",
              lat = lats,
              lon = lons,
              sdate = sdate,
              edate = edate,
              large = FALSE)

gc()

# Using cloud free reproceed data long term data set
# "https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/dataset-oc-glo-chl-multi_cci-l3-chl_4km_daily-rep-v02?"
# "https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/dataset-oc-glo-bio-multi-l4-chl_interpolated_4km_daily-rep?"

# Monthly
# url = "https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/dataset-oc-glo-bio-multi-l4-chl_4km_monthly-rep?"
# url = "https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/dataset-oc-glo-bio-multi-l4-chl_interpolated_4km_daily-rep?"

# chl = loadsat(var = "CHL",
#               sdate = sdate,
#               edate = edate,
#               lons = lons,
#               lats = lats,
#               org = "1900-01-01",
#               lat_varid = "lat",
#               lon_varid = "lon",
#               by = 30,
#               url = "https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/dataset-oc-glo-bio-multi-l4-chl_interpolated_4km_daily-rep?")

#url = "/thredds/dodsC/cmems_obs-oc_glo_bgc-plankton_my_l4-multi-4km_P1M"
```

correct oreantation
```{r, eval = FALSE}
chl = oreant(chl, t1 = TRUE)
```

Remove South of Hawaii and Chlorophyll front  
```{r, eval = TRUE, message = FALSE, warning = FALSE}
# chl = bufcoast(chl, 
#                region = "Hawaiian Islands", 
#                path = "../data/USMaritimeLimitsAndBoundariesSHP")

# I want all that gone
chl = polymask(chl)

gc()
```

Calculate CHL anomaly and boolian bloom array
```{r, message = FALSE, warning = FALSE}
# I receive an warning when detrending for cells of all NA values (Hawaii)
# chl anom. calc. needs at least one full year to work correctly
chl_a = anomalize(chl, detrend = TRUE)
chl_b = bool(chl_a)
gc()
```

Apply functions and catinate summary table
```{r, message = FALSE, warning = FALSE}
# I need to find the start and end index of each bloom
sumtbl = sumstats(blooms = chl_b, 
                  signal1 = chl, 
                  signal2 = chl_a, 
                  sdate = blooms$sdate, 
                  edate = blooms$edate)
gc()
```

saving tables as csv
```{r, eval = FALSE, message = FALSE, warning = FALSE,}
write.csv(sumtbl, file = "../data/full_sum_20230218.csv")
gc()
```

```{r, eval = FALSE, message = FALSE, warning = FALSE,}
sumtbl = read.csv("../data/full_sum_20230218.csv")
gc()
```


# Data visualization

## Summary table  

### summary table
```{r, message = FALSE, fig.show = "hold", out.width = "95%",  warning = FALSE}
tbl = sumtbl
rownames(tbl) = NULL
#idx = !year(tbl$sdate) == 2008
#tbl = tbl[idx,]
kbl(tbl, 
    booktabs = T, 
    latex_options = c("striped"),
    caption = "Summary statistics of each bloom around St. ALOHA and 30N between 2002 to 2019. Magnitude taken as the maximum CHL $mg/m^3$ value the blooms reached, and area is the maximum area the bloom reached.") %>%
  kable_styling(full_width = F, latex_options = "HOLD_position") %>%
  column_spec(1:2, bold = F, color = "black") %>%
  column_spec(1:7, width = "10em")
  # row_spec(c(4, 5, 8, 11, 12, 17), bold = F, color = "black", background = "#add8e6")
  # add_header_above(c("Table of (sub)mesoscale contributions: St. ALOHA" = 5), bold=T)
```

## Clim maps {.tabset}

### Raw CHL signal
```{r, message = FALSE, fig.show = "hold", out.width = "95%",  warning = FALSE}
csum <- subsum(chl)
cmap <- calc(csum, fun = mean, na.rm = TRUE)
plot(cmap)
```

### CHL anomaly signal
```{r, message = FALSE, fig.show = "hold", out.width = "95%",  warning = FALSE}
csum <- subsum(chl_a)
cmap <- calc(csum, fun = mean, na.rm = TRUE)
plot(cmap, main = "CHL anomaly")
```
### Raw CHL signal
```{r, message = FALSE, fig.show = "hold", out.width = "95%",  warning = FALSE}
csum <- subsum(chl)

tsa = cellStats(x = chl_a, stat = "mean", na.rm = TRUE)

plot(getZ(chl_a), tsa)
```

## Visualize Summary Table {.tabset}

### Big Bloom Magnitudes
```{r, message = FALSE, fig.show = "hold", fig.width = 10,  warning = FALSE}
tbl = sumtbl
tbl[y == 2010, ]$mag = 0.142
x = tbl$mag
y = year(tbl$sdate)

main = ""
ylab = TeX(r'(CHL $[mg/mg^3]$)')
xlab = "Bloom year"
# this changes the linewidth outsidet he plot
opar = par(lwd = 2)

par(mar = c(5,5,3,3))
barplot(x, 
        names.arg = y, 
        main = "",
        ylab = "", 
        xlab = "",
        ylim = c(0, max(x) + max(x)*0.05),
        axes = FALSE,
        cex.names = 1.25,
        col = "grey80",
        #las=2,
        cex.axis=0.2,
        bourder = colvect("grey22", alpha = 0.9)
        #density = 1
        )
box(which = "plot", lty = "solid", lwd = 3, col = colvect("grey22", alpha = 0.9))
grid(nx = 6, # X-axis divided in two sections
       ny = 6, # Y-axis divided in three sections
       lty = 2, 
       col = colvect(c("gray69"), alpha = 0.5), lwd = 1)
axis(side = 2,
     las  = 2,
     lwd  = 2,
     mgp  = c(1, 0.75, 0),
     cex.axis = 1.25,
     col = colvect("grey22", alpha = 0.9))
title(main = main,
      cex.lab = 2,
      line= 0.75,
      adj = 0)
title(ylab = ylab, cex.lab = 1.5, line = 2.75)
title(xlab = xlab, cex.lab = 1.5, line = 2.5)
```

### Big Bloom Areas
```{r, message = FALSE, fig.show = "hold", fig.width = 10,  warning = FALSE}
tbl = sumtbl
#idx = tbl$area > 550000
#tbl = tbl[idx,]
y = year(tbl$sdate)
tbl[y == 2022, ]$area = 682983.1
x = tbl$area/100000
y = year(tbl$sdate)
main = ""
ylab = TeX(r'(Area $[km \times 10^5]$)')
xlab = "Bloom year"

# this changes the linewidth outsidet he plot
par(mar = c(5,5,3,3))
opar = par(lwd = 2)
barplot(x, 
        axes = FALSE,
        names.arg = y, 
        main = "",
        ylab = "", 
        xlab = "",
        ylim = c(0, max(x) + max(x)*0.05))
box(which = "plot", lty = "solid", lwd = 3, col = colvect("grey22", alpha = 0.9))
grid(nx = 6, # X-axis divided in two sections
       ny = 6, # Y-axis divided in three sections
       lty = 2, 
       col = colvect(c("gray69"), alpha = 0.5), lwd = 1)
axis(side = 2,
     las  = 2, 
     lwd  = 2, 
     mgp  = c(1, 0.75, 0), 
     cex.axis = 1.25,
     col = colvect("grey22", alpha = 0.9))
title(main = main,
      cex.lab = 2,
      line= 0.75,
      adj = 0)
title(ylab = ylab, cex.lab = 1.5, line = 2.75)
title(xlab = xlab, cex.lab = 1.5, line = 2.5)
```

### Bloom Duration
```{r, message = FALSE, fig.width = 10, fig.height= 10, warning = FALSE}
tbl = sumtbl
#idx = tbl$area > 550000
#tbl = tbl[idx,]

start = yday(tbl$sdate)
end   = yday(tbl$edate)
# end[end < 100] = 365
# end[length(end)] = end[length(end)] + 60

durations = data.frame(id = as.character(year(tbl$sdate)), start, end)
durations$middle = apply(durations[, c("start", "end")], 1, FUN = mean, na.rm = TRUE)
durations$year = year(tbl$sdate)

# seperating years with two blooms to plot correctly
idx_dup   = duplicated(durations$year)
twobloom  = durations[idx_dup,]
durations = durations[!idx_dup,]

idx_dup    = duplicated(twobloom$year)
threebloom = twobloom[idx_dup,]
twobloom   = twobloom[!idx_dup,]

ggplot(durations) +
  geom_boxplot(
    fill = "grey100",
    colour = colvect("grey22", alpha = 0.9),
    size = 0.75,
    stat = "identity",
    aes(x = id, 
        lower  = start, 
        middle = middle, 
        upper  = end, 
        ymin   = start, 
        ymax   = end
        )) +
  geom_boxplot(data = twobloom,
    fill = "grey100",
    colour = colvect("grey22", alpha = 0.9),
    size = 0.75,
    stat = "identity",
    aes(x = id, 
        lower  = start, 
        middle = middle, 
        upper  = end, 
        ymin   = start, 
        ymax   = end
        )) +
  geom_boxplot(data = threebloom,
    fill = "grey100",
    colour = colvect("grey22", alpha = 0.9),
    size = 0.75,
    stat = "identity",
    aes(x = id, 
        lower  = start, 
        middle = middle, 
        upper  = end, 
        ymin   = start, 
        ymax   = end
        )) +
  xlab("Bloom year") +
  ylab("Day of year") +
  labs(title = "") + 
  theme_bw() + 
  theme(
    panel.border = element_rect(colour = colvect("grey22", alpha = 0.9), 
                                fill = NA, 
                                size = 2),
    axis.title.x = element_text(size = rel(1.25)),
    axis.title.y = element_text(size = rel(1.25)),
    axis.text.x = element_text(size = rel(1.25)),
    axis.text.y = element_text(size = rel(1.25)),
    plot.title = element_text(size  = rel(1.25))
    ) +
  coord_flip()
```

### Bloom Center's
```{r, message = FALSE, fig.width = 10, fig.height = 6.8, fig.show = "hold", out.width = "95%",  warning = FALSE}

tbl = sumtbl
e = extent(-173, -125, 11, 46)
wdmap <- getMap(resolution = "high")
temp = tbl
idx = tbl$lon > -165
tbl = tbl[idx,]
#idx = !year(tbl$sdate) == 2015
#tbl = tbl[idx,]

# radius in km
a = tbl$area
r = sqrt(a/pi) 
r = r/110.574 
tbl$rad = r 
tbl$y = as.character(year(tbl$sdate))

idx = order(tbl$area, decreasing = TRUE)
txt = tbl[idx,]
txt = txt[1:3,]

jamie_theme(e[1:2], e[3:4],
            main = "", 
            ylab = "Latitude", 
            xlab = "Longitude",
            dt = FALSE,
            xaxes = TRUE,
            yaxes = TRUE,
            mar = c(5,5,2,3)
            #asp = 1.08
            )
points(tbl$lon, tbl$lat, pch = 19)
for(i in 1:nrow(tbl)) draw.circle(tbl$lon[i], 
                                  tbl$lat[i], 
                                  tbl$rad[i],
                                  border = colvect("grey22", alpha = 0.6),
                                  lwd = 2,
                                  nv  = 500,
                                  lty = 2)
plot(wdmap, 
     xlim = e[1:2], 
     ylim = e[3:4], 
     asp = 1, 
     bg = "black", 
     border = "black", 
     col = "black", 
     #wrap = c(-180, 180), 
     add = TRUE)
text(txt$lon, txt$lat, label = txt$y, adj = 1.2)
box(which = "plot", lty = "solid", lwd = 3, col = colvect("grey22", alpha = 0.9))
#legend(-140, 22, legend = "centers", pch = 20)
```












# UNUSED CODE

STL Filter: calculate anomaly, seasonality and climatology 
```{r, message = FALSE, warning = FALSE}
# ts   = filt_stl(ras = chl, res = 1, sub.start = 8)
# gc()
```

Fill small blips and identify large blooms. Determine start, end, and duration of blooms
```{r, message = FALSE, warning = FALSE}
# durations = bloomspan(anom = ts$anom, time = getZ(chl), gaps = 1, dur = 1)
# gc()

# Create a list of objects subset by bloom time span to calc stats on
# tbl = sumstats(blooms = chl_b, signal = chl, tbl = durations)
```


FFT: calculate anomaly, seasonality and climatology 
```{r, eval = FALSE, message = FALSE, warning = FALSE}
fft_raw  = cellStats(chl, stat = mean, na.rm = TRUE)
fft_raw = as.numeric(fft_raw)
idx = !is.na(fft_raw)
fft_t = getZ(chl)
fft_t = fft_t[idx]
fft_raw = fft_raw[idx]

fft_clim = nff(x = fft_raw, n = 50L)
fft_clim = fft_clim$signal
fft_anom = fft_raw - fft_clim/100
gc()
```


## STL filter {.tabset}

### Raw
```{r, message = FALSE, fig.show = "hold", out.width = "75%",  warning = FALSE}
t = getZ(chl)
jamie_theme(t, ts$raw, main = "30N raw", dt = FALSE, ylim = range(ts$raw, na.rm = TRUE))
points(t, ts$raw, pch = 20)
```

### Anomaly
```{r, message = FALSE, fig.show = "hold", out.width = "75%",  warning = FALSE}
thresh = median(ts$anom, na.rm = TRUE) + mad(ts$anom, na.rm = TRUE)
idx    = ts$anom > thresh
anom_b = ts$anom[idx]
t_b = t[idx]

jamie_theme(t, ts$anom, main = "30N anomaly", dt = FALSE, ylim = range(ts$anom,na.rm = TRUE))
points(t, ts$anom, pch = 20)
points(t_b, anom_b, pch = 20, col = "red")

head(idx[100:200], 25)
```

### Climatology
```{r, message = FALSE, fig.show = "hold", out.width = "75%",  warning = FALSE}
jamie_theme(t, ts$clim, main = "30N climatology", dt = FALSE, ylim = range(ts$clim, na.rm = TRUE))
points(t, ts$clim, pch = 20) 
```

### Seasonanilty
```{r, message = FALSE, fig.show = "hold", out.width = "75%",  warning = FALSE}
jamie_theme(t, ts$seas, main = "30N seasonanilty", dt = FALSE, ylim = range(ts$seas))
points(t, ts$seas, pch = 20)
```

### Raw vs climate
```{r, message = FALSE, fig.show = "hold", out.width = "75%",  warning = FALSE}

jamie_theme(t, ts$raw, main = "aloha raw to climate comparison", dt = FALSE, ylim = range(ts$raw, na.rm = TRUE))
points(t, ts$raw, pch = 20)
points(t, ts$seas_clim, pch = 20, col = "red")
```

## FFT Data Vis {.tabset}

### Climate and raw signal
```{r, message = FALSE, fig.show = "hold", out.width = "75%",  warning = FALSE}
jamie_theme(fft_t, fft_raw, main = "aloha raw to climate comparison", dt = FALSE, ylim = range(fft_raw, na.rm = TRUE))
points(fft_t, fft_raw, pch = 20)
points(fft_t, fft_clim/100, pch = 20, col = "red")
```

### Anomaly signal
```{r, message = FALSE, fig.show = "hold", out.width = "75%",  warning = FALSE}
thresh = median(fft_anom, na.rm = TRUE) + mad(fft_anom, na.rm = TRUE)
idx    = fft_anom> thresh
fft_anom_b = fft_anom[idx]
fft_t_b = fft_t[idx]

jamie_theme(fft_t, fft_anom, main = "aloha raw to climate comparison", dt = FALSE, ylim = range(fft_anom, na.rm = TRUE))
points(fft_t, fft_anom, pch = 20)
points(fft_t_b, fft_anom_b, pch = 20, col = "red")
```














