---
title: "Case Study"
author: "Jamie Ash"
date: "05/31/2022"
output: 
  bookdown::html_document2:  
    self-contained: yes
    theme: paper #cerulean #cosmo #journal #readable
    toc: false
    smooth_scroll: true
    toc_depth: 3
    toc_float: true
    number_sections: false
    fig_caption: yes
    code_folding: hide
---  

<style type="text/css">  
/* Note: CSS uses C-style commenting. */
h1.title{font-size:22px; text-align:center;}
h4.author{font-size:16px; text-align:center;}
h4.date{font-size:16px; text-align:center;}
body{ /* Normal  */ font-size: 13px}
td {  /* Table   */ font-size: 12px}
h1 { /* Header 1 */ font-size: 16px}
h2 { /* Header 2 */ font-size: 14px}
h3 { /* Header 3 */ font-size: 12px}
.math{ font-size: 10pt;}
.hi{ /* hanging indents */ 
    padding-left:22px; 
    text-indent:-22px;
}
.main-container {
  max-width: 1000px;
  margin-left: auto;
  margin-right: auto;
}
blockquote {  
    padding: 10px 20px;
    margin: 0 0 20px;
    font-size: 12px;
    border-left: 5px solid #eee;
}
code.r{ /* code */ 
       font-size: 12px;
}
pre{/*preformatted text*/ 
    font-size: 12px;
}
p.caption {/* figure captions */ 
    font-size: 1.0em;
    font-style: italic; 
} 
.vscroll-plot {
    width: 800px;
    height: 400px;
    overflow-y: scroll;
    overflow-x: hidden;
}
body {
text-align: justify}
</style>

```{r setup, include=FALSE, message = FALSE}
source("functions.R")
source("libraries.R")
source("blooms.R")
rasterOptions(maxmemory = 120e+10, memfrac = 0.9)
knitr::opts_chunk$set(echo = TRUE)
```

Custom functions 
```{r, message = FALSE}
# input is a boolian raster layer of 0/1 
# output is a data frame with only high values included
# Downlsize reduces the size of the data set
# I may like to simultaniously index the chl signal as well
chl_fronts = function(blooms,
                      signal,
                      downsize = 10){
  ras  = raster::flip(blooms, direction = "x")

  idx  = which(values(blooms) == TRUE) # == TRUE
  
  e    = extent(blooms)
  s    = dim(blooms)
  lons = seq(from = e[1], to = e[2], length = s[1]) 
  lats = seq(from = e[3], to = e[4], length = s[2]) 
  grid = pracma::meshgrid(lons, lats)
  # using the blooms index to subset the signal values
  fronts = data.frame(lons = grid$X[idx], lats = grid$Y[idx], value = signal[idx])
  idx = seq(1, nrow(fronts), downsize)
  fronts = fronts[idx, ]
  fronts
}

# use the apply function to apply the calc mean function for every 8 layers

avestack = function(ras, by = 8, fun){
  t = getZ(ras)
  s      = dim(ras)
  idx_8d = rep(1:ceiling(s[3]/8), 8)
  idx_8d = sort(idx_8d)
  rem    = abs(s[3] - length(idx_8d))
  idx_8d = idx_8d[1:(length(idx_8d)-rem)]
  idx_og = 1:s[3]
  t_idx  = unique(idx_8d)
  idx    = data.frame(idx_8d = idx_8d, idx_og = idx_og)
  idx$idx_8d = as.factor(idx$idx_8d)
  idx = split(idx, idx_8d)
  
  # subset and average every eight days
  tempset = function(x, ras) ras[[x$idx_og]]
  days = lapply(idx, FUN = tempset, ras = ras)
  days = lapply(days, FUN = calc, fun = fun, na.rm = TRUE)
  days = brick(days)
  days = setZ(days, z = t[t_idx], name = "time")
  days
  }

# Writing this as a function to take a date range, a chl anomaly signal, 
# and output an area time-series

# Input is 3D chla raster output is time series vector of bloom area
bloom_area = function(b){
  t    = getZ(b)
  rr   = reclassify(b, cbind(0,NA))
  a    = raster::area(rr)
  temp = raster::mask(a, rr)
  
  km2  = cellStats(temp, stat = "sum", na.rm = TRUE)
  data.frame(area = as.numeric(km2), time= t)
}

# input x is a raster of raw chl
# inpur x is a raster of 0/1 bloom non-bloom
bloom_con = function(x, b, stat = "mean"){
  t = getZ(x)
  # make bloom raster a mask
  rr   = reclassify(b, cbind(0,NA))
  # mask bloom raster with bloom raster
  temp = raster::mask(x, rr)
  mag  = cellStats(temp, stat = stat, na.rm = TRUE)
  data.frame(con = as.numeric(mag), time= t)
}  

basic_theme <- function(x,
                        y, 
                        mar = c(1,1,1,1),
                        #asp = 1,
                        ylim = range(y),
                        xlim = range(x),
                        main = "",
                        dt = TRUE,
                        ylab = "",
                        xlab = ""
){
  if(is.numeric(x)) x <- round(x, 1)
  if(is.numeric(y)) y <- round(y, 1)
  par(mar = mar)
  plot(0,
       ylim = ylim,
       xlim = xlim,
       main = "",
       xlab = "",
       ylab = "",
       axes = FALSE,
       )
  grid(nx = NULL, # X-axis divided in two sections
       ny = NULL, # Y-axis divided in three sections
       lty = 2, 
       col = colvect(c("gray69"), alpha = 0.5), lwd = 1)
  box(which = "plot", 
      lty = "solid", 
      lwd = 3, 
      col = colvect("grey22", alpha = 0.9))
  title(main = main,
        cex.lab = 2,
        line= 0.75,
        adj = 0)
  title(ylab = ylab, cex.lab = 1.5, line = 2.5)
  title(xlab = xlab, cex.lab = 1.5, line = 2.5)
}


# 
# area_ts = function(x = chla, sdate = NULL, edate = NULL, stat = "sum") {
#   
#   # Find a bloom not bloom matrix
#   t = getZ(x)
#   u = calc(x, fun = median, na.rm = TRUE)
#   o = calc(x, fun = mad, na.rm = TRUE)
#   blooms = x > u + o
#   blooms = setZ(blooms, z = t, name = "time")
#   
#   # subset the bloom by time
#   blooms  <- crop3d(x     = blooms, 
#                     sdate = sdate,
#                     edate = edate)
#   
#   t <- getZ(blooms)
#   
#   # Area time series for duration 
#   rr <- reclassify(blooms, cbind(0,NA))
#   # find area of each cell + mask cells that matter
#   a  <- raster::area(rr)
#   aa <- raster::mask(a, rr)
#   # take the sum of all the cells area for bloom magnitude
#   area <- as.numeric(cellStats(aa, stat = "sum", na.rm = TRUE)) # m2 to km2?
#   
#   # I'm taking the sum of true/flase values. I want to mask the chla by the 
#   # true/false
#   mag  <- crop3d(x     = chla, 
#                  sdate = sdate,
#                  edate = edate)
#   
#   #mag <- raster::mask(mag, rr)
#   # I'm going to try masking this way so simple. 
#   # multiplying by boolian matrix of blooms
#   mag <- mag * blooms
#   val <- cellStats(mag, stat, na.rm = TRUE)
#   
#   data.frame(area = area, val = val, time = t)
#   # CHL timeseries. layer wise mean the seasonally filtered chl
# }

# polymask <- function(ras, 
#                      x = c(-170, -152, -157, -162, -170, -170), 
#                      y = c(  20,   20,   22,  24.5, 28, 20)){
#   Sr1 = Polygon(cbind(x,
#                       y))
#   Spp = SpatialPolygons(list(Polygons(list(Sr1), "s1")))
#   SpP_ras = raster::rasterize(Spp, ras, getCover=TRUE)
#   SpP_ras[SpP_ras == 1] = NA
#   ras <- raster::mask(ras, SpP_ras)
#   ras
# }
# input is a raster output is a boolian raster of bloom/not bloom

bool = function(x){
  u <- calc(x, fun = median, na.rm = TRUE)
  o <- calc(x, fun = mad, na.rm = TRUE)
  boo <- x > (u + o)
  extent(boo) <- extent(x)
  boo <- setZ(boo, z = getZ(x), name = "time")
  boo
}

vectorize  <- function(x) {
  sdate <- getZ(x)
  # x = raster::flip(x, direction = "x")
  # x = raster::flip(x, direction = "y")
  # 
  e = extent(x)
  x = t(x)
  extent(x) = e
  # 
  x     <- rasterToPoints(x)
  x     <- data.frame(x)
  colnames(x) <- c("lon", "lat", as.character(sdate))
  x     <- reshape2::melt(x, id.vars = c("lat", "lon"))
  colnames(x) <- c("lats", "lons", "time", "val")
  x
}

tsect <- function(x, y, z, xreach = 1, yreach = 1, xlen = 120, ylen = 40){
  #make a max min vector of sla and fsle
  coord = c(max(x, na.rm = TRUE), min(x, na.rm = TRUE), 
            max(y, na.rm = TRUE), min(y, na.rm = TRUE))
  
  # use maxmin vector to create fake x, y vectors (downsized) as meshgrid input
  y_vec <- seq(from = coord[4], to = coord[3], length.out = ylen) 
  # may need to invert to be same length 
  x_vec <- seq(from = coord[2], to = coord[1], length.out = xlen)
  xy_grid <- meshgrid(x_vec, y_vec)
  
  #inisilize a matrix of dim fs_grid[1] filled with NA values
  dims <- dim(xy_grid[[1]])
  z_grid <- matrix(data = NA, 
                   nrow = dims[1], 
                   ncol = dims[2], 
                   dimnames = NULL)
  
  for(iy in 1:dims[1]) {
    for(ix in 1:dims[2]){
      # where in the df is the difference greater than the units
      box_x <- which(abs(x - xy_grid[[1]][iy, ix]) <= xreach)
      box_y <- which(abs(y - xy_grid[[2]][iy, ix]) <= yreach)
      # I think the grids are dif sizes and should be subet differently
      #index vector of both cox_sla and box_fsle as one
      #box <- sort(match(box_y, box_x))
      
      # I think this is the correct way to do this
      box <- box_x[box_x %in% box_y]
      
      z_grid[iy, ix] <- mean(z[box], na.rm = TRUE)
    }
  }
  
  list(z_grid, y_vec, x_vec)
}

jamie_theme <- function(x,
                        y, 
                        line = 0.75,
                        adj = 0,
                        #asp = 1,
                        ylim = range(y),
                        xlim = range(x),
                        main = "",
                        #labels = seq(from = min(x)+5, to   = max(x)-5, by   = 15),
                        labels = round(seq(range(x)[1]+1, range(x)[2]-1, length = 6), 1),
                        at =  round(seq(range(x)[1]+1, range(x)[2]-1, length = 6), 1),
                        dt = TRUE,
                        yaxes = TRUE,
                        xaxes = TRUE,
                        ylab = "",
                        xlab = "",
                        cex.main = 1.25
){
  
  if(dt == TRUE) labels <- format(as.Date(labels), "%b-%d")
  if(is.numeric(x)) x <- round(x, 1)
  if(is.numeric(y)) y <- round(y, 1)
  
  plot(0,
       ylim = ylim,
       xlim = xlim,
       main = "",
       xlab = "",
       ylab = "",
       axes = FALSE,
       #asp = asp,
       )
  grid(nx = NULL, # X-axis divided in two sections
       ny = NULL, # Y-axis divided in three sections
       lty = 2, 
       col = colvect(c("gray69"), alpha = 0.5), lwd = 1)
  box(which = "plot", 
      lty = "solid", 
      lwd = 3, 
      col = colvect("grey22", alpha = 0.9))
  if(xaxes){axis(side = 1,
       at = at,
       labels = labels,
       las = 1, 
       lwd = 2, 
       mgp = c(2, 1, 0), 
       cex.axis = 1.15,
       col = colvect("grey22", alpha = 0.9))}
  if(yaxes){axis(side = 2,
       las  = 2, 
       lwd  = 2, 
       mgp  = c(1, 0.75, 0), 
       cex.axis = 1.15,
       col = colvect("grey22", alpha = 0.9))}
  title(main = main,
        cex.lab = cex.main,
        line = line,
        adj = adj)
  title(ylab = ylab, cex.lab = 1.25, line = 2.5)
  title(xlab = xlab, cex.lab = 1.25, line = 2.25)
}
```

# Data processing

# loading data via opendap had issues with over engineered function
```{r, warning = FALSE, message = FALSE}

sdate = as.Date("2018-01-01")
edate = as.Date("2019-01-01") 

# lons  = c(-158, -130)
# lats  = c(23, 34)
lons = c(-165, -130)
lats = c( 18,   35)
e = extent(lons, lats)

# cache_list()
cache_delete_all(force = FALSE)

# Set variables
lat_varid = "lat"
lon_varid = "lon"
# sdate = as.Date("1997-06-01")
# edate = as.Date("2021-11-01")
var = "CHL"
url = "https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/cmems_obs-oc_glo_bgc-plankton_my_l3-multi-4km_P1D?"

origin = "1900-01-01"

data = nc_open(url, verbose = FALSE, write = FALSE)
lat  = ncvar_get(data, varid = lat_varid)
lon  = ncvar_get(data, varid = lon_varid)
time = ncvar_get(data, varid = "time")
time = as.Date(time, origin = origin)

idx_lat  = which(lat > lats[1] & lat < lats[2])
idx_lon  = which(lon > lons[1] & lon < lons[2])
idx_time = which(time >= sdate & time <= edate)

idx_ras = paste("CHL",
                 paste("[", range(idx_time)[1], ":1:", range(idx_time)[2], "]", sep = ""),
                 paste("[", range(idx_lat)[1],  ":1:", range(idx_lat)[2],  "]", sep = ""),
                 paste("[", range(idx_lon)[1],  ":1:", range(idx_lon)[2],  "]", sep = ""),
                 sep = "")

idx_time = paste("time", paste("[", range(idx_time)[1], ":1:",range(idx_time)[2], "]", sep = ""), sep = "")
idx_lat  = paste(lat_varid, paste("[", range(idx_lat)[1], ":1:",range(idx_lat)[2],  "]", sep = ""), sep = "")
idx_lon  = paste(lon_varid, paste("[", range(idx_lon)[1], ":1:",range(idx_lon)[2],  "]", sep = ""), sep = "")
idx = paste(idx_lat, idx_lon, idx_time, idx_ras, sep = ",")

url = paste(url, idx, sep = "")

nc_close(data)
rm(data)

data = nc_open(url, verbose = FALSE, write = FALSE)

lat  = ncvar_get(data, varid = lat_varid)
lon  = ncvar_get(data, varid = lon_varid)
time = ncvar_get(data, varid = "time")
time = as.Date(time, origin = origin)
ras  = ncvar_get(data)
nc_close(data)

s   = dim(ras)
ras = raster::brick(ras)
ras = t(ras)
ras = setZ(ras, z = as.Date(time, origin = org), name = "time")
extent(ras) = extent(min(lons), max(lons), min(lats), max(lats))

chl = ras
rm(ras)
gc()
```

Loading the data using ERDDAP
```{r data, message = FALSE, warning = FALSE}
# Bloom start and stop dates
# blooms = read.csv("../data/outfiles/full_sum.csv")
# subseting just 2018 for case study
# idx = year(blooms$sdate) == 2018
# sdate = as.Date(blooms$sdate[idx]) - 16
# edate = as.Date(blooms$edate[idx]) + 15
# # lons  = c(-158, -130)
# # lats  = c(23, 34)
# lons = c(-165, -130)
# lats = c( 18,   35)
# 
# # cache_list()
# cache_delete_all(force = FALSE)
# 
# # should try loading chla data from ERDDAP. Or just anomaly this in one function 
# # I need to make sure I'm not saving a fuck tone of files
# # I'm uploading the large data set for the anomaly calculation
# chl = loadsat(url = "https://coastwatch.noaa.gov//erddap",
#               id  = "noaacwNPPN20S3ASCIDINEOFDaily",
#               lat = lats,
#               lon = lons,
#               parameter = c("all"),
#               sdate = as.Date("2018-02-09"),
#               edate = as.Date("2022-01-01"),
#               large = TRUE,
#               by = 10)

# ppp = loadsat(url = "https://coastwatch.pfeg.noaa.gov/erddap",
#               id  = "erdMH1pp8day",
#               lat = lats,
#               lon = lons,
#               parameter = c("productivity"),
#               sdate = as.Date("2018-06-01"),
#               edate = as.Date("2018-11-15"),
#               large = FALSE,
#               by = 10)

fields =  c("temp",
            "temp_qc",
            "temp_adjusted", 
            "psal", 
            "psal_adjusted", 
            "pres", 
            "pres_adjusted", 
            "time",
            "cycle_number", 
            "latitude", 
            "longitude",
            "platform_number")

float = loadfloat(sdate  = sdate, 
                  edate = edate, 
                  lons  = lons, 
                  lats  = lats,
                  fields = fields,
                  id = "ArgoFloats")
gc()
```

Correct oreantation
```{r}
# chl = oreant(chl, t1 = TRUE)
```

Remove Region around Hawaii. CZ fornow 
```{r, eval = TRUE, message = FALSE, warning = FALSE, cache = FALSE}
chl = bufcoast(chl, 
               region = "Hawaiian Islands", 
               path = "../data/USMaritimeLimitsAndBoundariesSHP")
gc()
```

Calculate CHL anomaly
```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE}
# I receive an warning when detrending for cells of all NA values (hawaii)
# chl anom. calc. needs at least one full year to work correctly
chla = anomalize(chl, detrend = FALSE)
chlb = bool(chla)
gc()
```

checking how well the filter did
```{r}
c_time = cellStats(chla, stat='mean', na.rm=TRUE)
plot(1:length(c_time), c_time)
```

Subset time-span of interest: 2018 summer bloom
```{r, message = FALSE, warning = FALSE, cache = FALSE}
# I could also just load using errdap. This is faster tho
#ppp_2018  = timesnip(ppp,  as.Date(sdate), as.Date(edate)
idx = year(blooms$sdate) == 2018
sdate = as.Date(blooms$sdate[idx]) - 16
edate = as.Date(blooms$edate[idx]) + 15

chl_2018  = timesnip(chl,  as.Date(sdate), as.Date(edate))
chla_2018 = timesnip(chla, as.Date(sdate), as.Date(edate))
chlb_2018 = timesnip(chlb, as.Date(sdate), as.Date(edate))
```

Process PP: resize ppp to chl and interpolate over clouds
```{r, eval = FALSE}
# matches x to y in the time domain
# match = function(x, y){
#   ty = getZ(y)
#   ty = as.numeric(ty)
#   tx = getZ(x)
#   tx = as.numeric(tx)
#   
#   idx  = which(is.element(tx, ty))
#   tx   = subset(tx, is.element(tx, ty))
#   x = raster::subset(x, idx)
#   x = setZ(x, z = as.Date(tx), name = "time")
#   x
# }
# resample = function(ras, to, method = "bilinear"){
#   raster::crs(ras) = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" 
#   t = getZ(to)
#   e = extent(to)
#   extent(ras) = extent(e)
#   ras = raster::resample(ras,  to, method = method)
#   extent(ras) = e
#   ras = setZ(ras, z = as.Date(t), name = "time") 
#   ras
#   }
# 
# time = getZ(ppp_2018)
# 
# chlb_8day = match(chlb_2018, ppp_2018)
# ppp_2018 = crop(ppp_2018, chlb_8day)
# extent(ppp_2018) = extent(chlb_8day)
# 
# ppp_2018 = resample(ppp_2018, to = chlb_8day)
# 
# # interpolate bitch
# ppp_2018 = approxNA(ppp_2018, rule=2)
# 
# # remove coastal effects
# ppp_2018 = bufcoast(ppp_2018, 
#                     region = "Hawaiian Islands", 
#                     path = "../data/USMaritimeLimitsAndBoundariesSHP")
# 
# ppp_2018 = setZ(ppp_2018, z = time, name = "time")
# extent(ppp_2018) = extent(chlb_8day)
# gc()
```

Calculate patch area
```{r, message = FALSE}
# integrated = depth integrated
tsa = bloom_area(chlb_2018)
# magnitude might like the raw CHL
tsc = bloom_con(chl_2018, chlb_2018, stat = "mean")

# Need to have a chlb_2018 that's 8 day time snipped
# calcuates the average pp might want total
#tsp = bloom_mag(ppp_2018, chlb_8day)

#tsa = area_ts(chla, sdate = sdate, edate = edate, stat = "mean")
bloom_2018 = data.frame(area = tsa$area, 
                        con  = tsc$con, 
                        time = tsa$time)

# cleaning work space
rm(tsa, tsc)
```

Process float data
```{r, message = FALSE, warning = FALSE}
# time-series of the mixed layer depth for the 
# Converting all columns to numeric
float$time = as.numeric(as.Date(float$time))

#float$float_serial_no = as.factor(float$float_serial_no)
float$float_serial_no = as.factor(float$platform_number)

float = data.frame(lapply(float, function(x) as.numeric(as.character(x))))

# removing bad float data
float = subset(float, 
               psal > 3 
               & temp > 5 
               & temp < 45 
               & pres > 0
               & pres < 800)

#float$float_serial_no <- as.numeric(float$float_serial_no)
float$float_serial_no <- as.numeric(float$platform_number)
float$time            <- as.numeric(float$time)

# Identifying bad floats and removing them
badf  = subset(float, temp < 10 & pres < 100)
#badid = is.element(float$float_serial_no, badf$float_serial_no)
badid = is.element(float$float_serial_no, badf$platform_number)
badcy = is.element(float$cycle_number, badf$cycle_number)
float = subset(float, !badid & !badcy)
rm(badcy, badid, badf)

# Identifying bad floats and removing them
badf = subset(float, temp > 20 & pres > 300)
#badid = is.element(float$float_serial_no, badf$float_serial_no)
badid = is.element(float$float_serial_no, badf$platform_number)
badcy = is.element(float$cycle_number, badf$cycle_number)
float = subset(float, !badid & !badcy)

rm(badcy, badid, badf)
```

Calculate density and mixed layer depth
```{r, message = FALSE}
# I changed the region so floats suck

# Density Calculation: function from the oce package and calc. density
float$rho = swRho(salinity = float$psal,
                  temperature = float$temp,
                  pressure = float$pres,
                  latitude = float$latitude,
                  eos = getOption("oceEOS", default = "gsw"))
# I should make sure float factors are not lost to NA values

#float = subset(float, !is.na(float_serial_no))
float = subset(float, !is.na(platform_number))

# Mixed layer depth per profile
#id = unique(float[, c("float_serial_no", "cycle_number")])
id = unique(float[, c("platform_number", "cycle_number")])

mixed = data.frame(matrix(data = NA, 
                          nrow = nrow(id), 
                          ncol = ncol(float)+1, 
                          dimnames = NULL))
colnames(mixed) = colnames(float)

for(i in 1:nrow(id)){
  oneprof = subset(float, 
                   float_serial_no == id$platform_number[i] & 
                   cycle_number == id$cycle_number[i])
  if(nrow(oneprof) > 4) mixed[i, ] = mldchu(ctd = oneprof, x = "rho", y = "pres_adjusted", n = 10)
}
```

Last minute remove bad data
```{r, message = FALSE}
mix = mixed
mix = subset(mixed, !pres == 8)
mix = subset(mix, !pres > 100)
chl_float = subset(float, pres < 40) 
```

Loess running average
```{r, message = FALSE}
# I may want to change tsa to include magnitude. Maybe keep seperate
t     = as.numeric(bloom_2018$time)

curve = loess(mix$pres_adjusted ~ mix$time) 
x_mod = seq(min(t), max(t), length.out = length(t))
mix_mod = predict(curve, newdata =  x_mod, type = "response")

bloom_2018$mld = mix_mod
```

CHL concentration and units conversion
```{r, message = FALSE}
# quik and dirty way of handling clouds. I hope hawiann islands are not in this extent
s = dim(chl_2018)
n = s[1] * s[2]
clouds = sum(is.na(values(chl_2018))) / n
clouds = unname(clouds)
bloom_2018$area =  bloom_2018$area * (1+clouds)

# In mg
bloom_2018$bio = (bloom_2018$mld * bloom_2018$con * bloom_2018$area * 1e+06)
# mg to kilotons 
bloom_2018$bio = bloom_2018$bio/(1e+12)
bloom_2018$area = bloom_2018$area/1e+05 #10e05
bloom_2018$con = bloom_2018$con*1000 #to ug
elephants = max(bloom_2018$bio*1000)/ 5.98742 # elephant in t
```

# Data visualization 

```{r, fig.width=5, fig.height=8, message = FALSE, fig.show = "hold",  warning = FALSE, echo = FALSE}
png(filename="biomass_20230331.png",
    width = 3.5,
    height = 6,
    units = "in",
    res = 300,
    pointsize = 10)

width = 3
width_px = 3 * 300
cex = width

layout(matrix(c(1,  
                2,
                3), 
              nrow = 3, 
              ncol = 1, 
              byrow = TRUE),
       widths  = c(1),
       heights = c(1, 1, 1.15))

# Area
par(mar = c(0, 5, 2, 3))
x = as.numeric(bloom_2018$time)
y = bloom_2018$area
jamie_theme(x = x, 
            y = y, 
            dt  =FALSE,
            main = "", 
            xlab = "",
            #ylim = c(2, 25),
            #xlim = c(as.Date("2018-07-01"), as.Date("2018-11-05")),
            ylab = expression(Surface ~ Area ~ (km^2 ~ e05)),
            line = 0.75,
            adj = 0,
            yaxes = TRUE,
            xaxes = FALSE)
points(x, y,  pch = 20, cex = 1.25) 
box(which = "plot", 
      lty = "solid", 
      lwd = 1, 
      col = colvect("grey22", alpha = 0.9))

# CHL Concentration
par(mar = c(1, 5, 1, 3))
x = bloom_2018$time
y = bloom_2018$bio
jamie_theme(x = x,
            y = y,
            #xlim = c(as.Date("2018-07-01"), as.Date("2018-11-05")),
            #ylim = c(0, 10),
            dt = FALSE,
            line = 0.75,
            adj = 0.0,
            main = "",
            xlab = "",
            yaxes = TRUE,
            ylab = expression(Chlorophyll ~ Biomass ~ (kt)),
            xaxes = FALSE)
points(x, y, pch = 20, cex = 1.25)
box(which = "plot", 
      lty = "solid", 
      lwd = 1, 
      col = colvect("grey22", alpha = 0.9))

# CHL Summed Magnitude
par(mar = c(4, 5, 0, 3))
y = bloom_2018$con/1000
x = as.Date(as.numeric(bloom_2018$time))
jamie_theme(x = x, 
            y = y, 
            #xlim = c(as.Date("2018-07-01"), as.Date("2018-11-05")),
            dt = TRUE,
            ylim = c(60, 145),
            line = -1,
            adj = 0.03,
            main = "", 
            xaxes = TRUE,
            yaxes = TRUE,
            xlab = "Date Time",
            ylab = expression(Chlorophyll ~ (ug ~ m^{-3}))
            )
points(x, y,  pch = 20, cex = 1.25)
box(which = "plot", 
      lty = "solid", 
      lwd = 1, 
      col = colvect("grey22", alpha = 0.9))
#dev.off()
```

```{r, fig.width=8, fig.height=5}
wdmap = getMap(resolution = "high")

x = -150
y = 27
a = 23*1e05
r = sqrt(a/pi) 
r = r/110.574 #latitude to meters

plot(x, y, 
     ylim = c(17, 37), 
     xlim = c(-170, -125),
     pch = 19,
     ylab = "Latitude",
     xlab = "Longitude")
draw.circle(x, 
            y, 
            r,
            border = colvect("grey22", alpha = 0.6),
            lwd = 2,
            nv  = 500,
            lty = 2)
plot(wdmap, 
     xlim = e[1:2], 
     ylim = e[3:4], 
     asp = 1, 
     bg = "black", 
     border = "black", 
     col = "black", 
     #wrap = c(-180, 180), 
     add = TRUE)
grid(nx = NULL, # X-axis divided in two sections
     ny = NULL, # Y-axis divided in three sections
     lty = 2, 
     col = colvect(c("gray69"), alpha = 0.5), lwd = 1)
box(which = "plot", lty = "solid", lwd = 3, col = colvect("grey22", alpha = 0.9))
```


### Mixed Layer Depth
```{r, message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE}
# Mixed Layer Depth
curve = loess(mix$pres_adjusted ~ as.numeric(mix$time)) 
curve = predict(curve, type = "response")

jamie_theme(x = mix$time, y = mix$pres_adjusted, 
            ylim = c(80, 0),
            main = "2018 Bloom Mixed Layer Depth", 
            xlab = "Date Time",
            ylab = "Pressure [dbar]")
points(mix$time, mix$pres_adjusted, pch = 20, cex = 0.6)
points(mix$time, curve, pch = 20, cex = 1)
points(bloom_2018$time, bloom_2018$mld, col = "red", lwd = 2, pch = 20)
```

## Argo-floats tracks  
```{r,  fig.height=8, message = FALSE, fig.show = "hold",  warning = FALSE}
# Float Tracks
e = extent(-165, -132, 18, 37)
wdmap = getMap(resolution = "high")

#id <- unique(float[, c("longitude", "latitude", "float_serial_no", "time")])
id <- unique(float[, c("longitude", "latitude", "platform_number", "time")])
#id$float_serial_no <- as.factor(id$float_serial_no)
id$float_serial_no <- as.factor(id$platform_number)
#nb.cols <- length(unique(id$float_serial_no))
nb.cols <- length(unique(id$platform_number))
mycolors <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)
#levels(id$float_serial_no) <- mycolors
levels(id$platform_number) <- mycolors

jamie_theme(x = id$longitude, y = id$latitude, 
            dt = FALSE,
            main = "2018 bloom argo-floats tracks", 
            xlab = "Longitude",
            ylab = "Latitude",
            xlim = e[1:2], 
            ylim = e[3:4])
sz <- scale01(id$time)
points(id$longitude, id$latitude,
       #col = as.character(id$float_serial_no),
       col = as.character(id$platform_number),
       cex = sz*1.5,
       pch = 20) 
plot(wdmap, 
     xlim = e[1:2], 
     ylim = e[3:4], 
     asp = 1, 
     bg = "black", 
     border = "black", 
     col = "black", 
     #wrap = c(-180, 180), 
     add = TRUE)
box(which = "plot", lty = "solid", lwd = 3, col = colvect("grey22", alpha = 0.9))
```





