---
title: "Untitled"
author: "Jamie Ash"
date: "2022-10-28"
output: pdf_document
---

I'm still using pre-downloaded data here... 

```{r setup, include=FALSE, message = FALSE}
source("functions.R")
source("libraries.R")
rasterOptions(maxmemory = 120e+10, memfrac = 0.9)
knitr::opts_chunk$set(echo = TRUE)
```

setting variables
```{r}
#blooms = read.csv("../data/outfiles/full_sum.csv")
# subseting just 2018 for case study
idx = year(blooms$sdate) == 2018
sdate = as.Date(blooms$sdate[idx]) - 22
edate = as.Date(blooms$edate[idx]) + 22
# lons  = c(-158, -130)
# lats  = c(23, 34)
lons = c(-171, -129)
lats = c( 17,   36)
e = extent(-165, -130, 20, 33)
```

functions
```{r}
# input is a boolian raster layer of 0/1 
# output is a data frame with only high values included
# Downsize reduces the size of the data set
# I may like to simultaneously index the chl signal as well

# ------------------------------------------------------------------------------
# input is a raster output is a boolian raster of bloom/not bloom
bool = function(x){
  u <- calc(x, fun = median, na.rm = TRUE)
  o <- calc(x, fun = mad, na.rm = TRUE)
  boo <- x > (u + o)
  extent(boo) <- extent(x)
  boo <- setZ(boo, z = getZ(x), name = "time")
  boo
}

# ---------------------------------------------------------------------------
tableit <- function(x) {
  vectorize  <- function(x) {
    sdate <- getZ(x)
    x     <- rasterToPoints(x)
    x     <- data.frame(x)
    colnames(x) <- c("lat", "lon", as.character(sdate))
    x     <- reshape2::melt(x, id.vars = c("lat", "lon"))
    colnames(x) <- c("lats", "lons", "time", "val")
    x
  }
  x     = lapply(x, FUN = vectorize)
  vals  = lapply(x, FUN = subset, select = "val")
  quord = lapply(x, FUN = subset, select = c("lons", "lats", "time"))
  xyz   = data.frame(quord[1], vals)
  
  xyz$time = as.character(xyz$time)
  xyz$time = as.Date(xyz$time)
  xyz
  
  # s = dim(x)
  # grid = meshgrid(lon, lat)
  # lat  = raster(grid$X)
  # lat  = raster(grid$Y)
  # lat  = rep(lat, s[3])
  # lon  = rep(lon, s[3])
  # lat  = values(raster(lat))
  # lon  = values(raster(lat))
}

# ------------------------------------------------------------------------------
jamie_theme <- function(x,
                        y, 
                        mar = c(1,1,1,1),
                        #asp = 1,
                        ylim = range(y),
                        xlim = range(x),
                        main = "",
                        dt = TRUE,
                        ylab = "",
                        xlab = "",
                        yaxes = FALSE,
                        xaxes = FALSE,
                        line = 0.75,
                        adj = 0.1
                        ){
  if(is.numeric(x)) x <- round(x, 1)
  if(is.numeric(y)) y <- round(y, 1)
  par(mar = mar)
  plot(0,
       ylim = ylim,
       xlim = xlim,
       main = "",
       xlab = "",
       ylab = "",
       axes = FALSE,
       )
  if(xaxes){
    axis(side = 1,
         las = 1, 
         lwd = 2, 
         mgp = c(2, 1, 0), 
         cex.axis = 1,
         col = colvect("grey22", alpha = 0.9))
    }
  if(yaxes){
    axis(side = 2,
         las  = 2, 
         lwd  = 2, 
         mgp  = c(1, 0.75, 0), 
         cex.axis = 1,
         col = colvect("grey22", alpha = 0.9))
    }
  grid(nx = NULL, # X-axis divided in two sections
       ny = NULL, # Y-axis divided in three sections
       lty = 2, 
       col = colvect(c("gray69"), alpha = 0.5), lwd = 1)
  box(which = "plot", 
      lty = "solid", 
      lwd = 3, 
      col = colvect("grey22", alpha = 0.9))
  title(main = main,
        cex.lab = 2,
        line= line,
        adj = adj)
  title(ylab = ylab, cex.lab = 1.5, line = 2.5)
  title(xlab = xlab, cex.lab = 1.5, line = 2.5)
}

```

load data
```{r}
chl = load_nc(path = "..\\data\\chl\\",
              patt = "chl_20180624_20181118.nc",
              origin = "1970-01-01",
              vars = c("lats", "lons", "value","time"))

# I saved chl incorrectly
c_e = extent(chl)
extent(chl) = extent(c_e[3], c_e[4], c_e[1], c_e[2])
rm(c_e)

fsle = load_nc(path = "C:\\Users\\james\\Desktop\\jamieslife\\data\\infiles\\case_study\\fsle\\",
               patt = "fsle_po_2018.nc")
extent(fsle) = extent(fsle) - c(360, 360, 0 ,0)
gc()

sla = load_nc(path = "C:\\Users\\james\\Desktop\\jamieslife\\data\\infiles\\case_study\\sla\\",
              patt = "2010_2021_sla.nc",
              vars = c("latitude", "longitude", "sla", "time"))
gc()
```

fix oreantation
```{r}
fsle = oreant(fsle, flip = "y", t1 = TRUE)
gc()
sla  = oreant(sla, flip = "y", t1 = TRUE)
gc()
```

Crop all to the same extent
```{r crop}
e = extent(-158, -130, 22, 33)
fsle = raster::crop(fsle, e)
chl  = raster::crop(chl, e)
sla  = raster::crop(sla,  e)
gc()
```

Calculate a sla anomaly
```{r}
slaa = anomalize(ras = sla, detrend = FALSE)
rm(sla)
gc()
```

Make FSLE and CHL them the same time domain. Make FSLE same temporal resolution as CHL
```{r}
match = function(x, y){
  ty = getZ(y)
  ty = as.numeric(ty)
  tx = getZ(x)
  tx = as.numeric(tx)
  
  idx  = which(is.element(tx, ty))
  tx   = subset(tx, is.element(tx, ty))
  x = raster::subset(x, idx)
  x = setZ(x, z = as.Date(tx), name = "time")
  x
}

fsle = match(x = fsle, y = chl)
slaa = match(x = slaa, y = chl)
gc()
```

Cut to timespan of bloom from table 1
Not completly nessisary now will see for latter
```{r, eval = FALSE}
edate = edate - 11
chl  = timesnip(chl,   sdate, edate)
fsle = timesnip(fsle,  sdate, edate)
slaa = timesnip(slaa,  sdate, edate)
```

Looking at oreantation and projections there's a huge projection issue
```{r}
fmap  = calc(fsle, fun = mean)
smap  = calc(slaa,  fun = mean)
wdmap = getMap(resolution = "high")

e = extent(fmap)
plot(fmap)
plot(wdmap, 
       xlim = e[1:2], 
       ylim = e[3:4], 
       asp = 1, 
       bg = "white", 
       border = "black", 
       col = "ivory", 
       #wrap = c(-180, 180), 
       add = TRUE)

e = extent(smap)
plot(smap)
plot(wdmap, 
       xlim = e[1:2], 
       ylim = e[3:4], 
       asp = 1, 
       bg = "white", 
       border = "black", 
       col = "ivory", 
       #wrap = c(-180, 180), 
       add = TRUE)

e = extent(chl[[10]])
plot(chl[[10]])
plot(wdmap, 
       xlim = e[1:2], 
       ylim = e[3:4], 
       asp = 1, 
       bg = "white", 
       border = "black", 
       col = "ivory", 
       #wrap = c(-180, 180), 
       add = TRUE)
```

Resample to the size of FSLE
```{r resample}
resample = function(ras, to, method = "bilinear"){
  raster::crs(ras) = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" 
  t = getZ(to)
  e = extent(to)
  extent(ras) = extent(e)
  ras = raster::resample(ras,  to, method = method)
  extent(ras) = e
  ras = setZ(ras, z = as.Date(t), name = "time") 
  ras
  }

fsle = resample(fsle, to = chl, method = "ngb")
gc()
slaa = resample(slaa, to = chl, method = "ngb")
gc()
chl  = resample(chl,  to = chl, method = "ngb")
```

Remove coastal influence: SOmething is wrong here
```{r}
chl = bufcoast(chl, 
               region = "Hawaiian Islands", 
               path = "../data/USMaritimeLimitsAndBoundariesSHP")
fsle = bufcoast(fsle, 
               region = "Hawaiian Islands", 
               path = "../data/USMaritimeLimitsAndBoundariesSHP")
slaa = bufcoast(slaa, 
               region = "Hawaiian Islands", 
               path = "../data/USMaritimeLimitsAndBoundariesSHP")
gc()
```

Create table
```{r}
# this is much simpler code but may take alot of memory
vectorize = function (ras) {
  df = cbind(coordinates(ras), values(ras))
  df = data.frame(df)
  time = getZ(ras)
  time = rep(time, times = nrow(df))
  df$time = time
  colnames(df) = c("lon", "lat", "value", "time")
  df
  }

ras2tbl = function(ras, name = "value"){
  ras_l = list()
  for(i in 1:dim(ras)[3]) ras_l[[i]] = ras[[i]]
  
  ras_df = lapply(ras_l, vectorize)
  
  ras_df = do.call(rbind, ras_df)
  ras_df
  }

slaa_df = ras2tbl(slaa, name = "value")
gc()
fsle_df = ras2tbl(fsle, name = "value")
gc()
chl_df  = ras2tbl(chl,  name = "value")


# create one data frame
xyz = data.frame(time = slaa_df$time, 
                 lat  = slaa_df$lat, 
                 lon  = slaa_df$lon, 
                 fsle = fsle_df$value,
                 slaa = slaa_df$value,
                 chl  = chl_df$value)

rm(slaa_df, fsle_df, chl_df)
gc()
```

K-cluster by day produce table
```{r, message = FALSE, eval = FALSE}
# changing sla to slaa funciton should be written better
kcluster = function(x, y){
  y = scale01(y)
  x = scale01(x)
  
  us = median(x, na.rm = TRUE)
  uf = median(y, na.rm = TRUE)
  ms = mad(x, na.rm = TRUE)
  mf = mad(y, na.rm = TRUE)
  
  cents = rbind(c(us,      uf),
                c(us - ms, uf),
                c(us + ms, uf),
                c(us, uf + mf))
  set.seed(225)
  ktbl = kmeans(data.frame(x, y), # think it needs to be a data frame
                centers   = cents,
                iter.max  = 10,
                nstart    = 20, 
                algorithm = "Lloyd")
  
  k = data.frame(ktbl$cluster)[, 1]
  as.numeric(k)
}

# k-means does not like NA values
xyz = subset(xyz, !is.na(fsle))
xyz = subset(xyz, !is.na(slaa))

# apply to the entire dataset
k = kcluster(x = xyz$slaa, y = xyz$fsle)
k = as.numeric(k)
xyz$k = k
gc()
```

Preparing a data frame for the GAM
```{r}
# remove effects of clouds
df = xyz

# Cleaning up the data set
df  = xyz[!is.na(df$chl),]
df$time = as.Date(df$time)
df$regions = as.factor(df$k)
gc()
# Converting lat lon to distance
xy   = cbind(df$lat, df$lon)
utms = rgdal::project(xy, "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0") # need to change zones
df$northing = utms[, 1]/1000
df$easting  = utms[, 2]/1000
df$doy = as.numeric(format(df$time, '%j'))
df = df[, c("chl", "slaa", "fsle", "regions", "doy", "easting", "northing")]
df$fsle = df$fsle * -1

# cutting the end date back
idx = df$doy < 309
df = df[idx, ]

gc()
```

perform GAM 
```{r}
library(rgdal)
library(mgcViz)
# gam_2018 = gam(chl ~ s(slaa) + s(fsle) + s(doy) + s(easting, northing), 
#                method = "REML",
#                data = df, 
#                family = gaussian(link = "identity"))

gam_2018 = gam(chl ~ s(slaa) + s(fsle) + s(doy) + s(easting, northing), 
               #method = "REML",
               data = df, 
               family = Gamma(link = "identity"))

# gam_2018 = gam(chl ~ s(slaa) + s(fsle) + s(doy), # + s(easting, northing), 
#                # method = "REML",
#                data = df, 
#                family = gaussian(link = "identity"))
# 
# 
# gam_2018 = gam(chl ~ s(doy, by = regions, k=4) + s(doy), #+ s(easting, northing), 
#                # method = "REML",
#                data = df, 
#                family = gaussian(link = "identity"))

summary(gam_2018)
```

Initial plot
```{r}
b = plot(gam_2018, all = TRUE)
```

Visualizing the gam
```{r, message = FALSE, fig.show = "hold",  fig.align = "center", fig.height = 4, fig.width = 12, fig.height = 5,  warning = FALSE}
start = as.Date("2018-06-15")
end   = as.Date("2018-08-25")
title = c("a)", "b)", "Time")

x = as.data.frame(cbind(b[[1]]$x, b[[1]]$fit, b[[1]]$se))
y = as.data.frame(cbind(b[[2]]$x, b[[2]]$fit, b[[2]]$se))
z = as.data.frame(cbind(b[[3]]$x, b[[3]]$fit, b[[3]]$se))

layout.matrix = matrix(c(1, 
                         2), 
                        nrow = 2,
                        ncol = 2,
                        byrow = TRUE)
nf = graphics::layout(layout.matrix, 
                       heights = c(2, 2))

shadedline(x    = x$V1, 
           ylim = c(-0.007, 0.01),
           xlim = c(-0.25, 0.2),
           y0 = x$V2,
           y1 = x$V2 + x$V3, 
           y2 = x$V2 - x$V3,
           title = "",
           line = -1.5,
           adj  = 0.02,
           main = paste(title[1]),
           cex.main = 1.25,
           ylab = "CHL Effect [mg/m^3]",
           col  = "grey69",
           xlab = "SLA",
           xaxis = FALSE,
           #labels = format(as.Date(x$V1), "%Y-%m-%d"),
           mar = c(5, 5, 1, 1)
           )
axis(side = 1, 
     las = 1, 
     lwd = 2, 
     mgp = c(2, 1, 0), 
     cex.axis = 1.15)

shadedline(x = y$V1, 
           ylim = c(min(y$V2 - y$V3), max(y$V2 + y$V3)),
           # xlim = c(0, ),
           xaxis = FALSE,
           y0 = y$V2,
           y1 = y$V2 + y$V3, 
           y2 = y$V2 - y$V3,
           title = "",
           line = -1.5,
           adj = 0.02,
           main = paste(title[2]),
           cex.main = 1.25,
           ylab = "",
           col = "grey69",
           xlab = "FSLE",
           #labels = format(as.Date(y$V1), "%Y-%m-%d"),
           mar = c(5, 3, 1, 3)
           )
axis(side = 1, 
     las = 1, 
     lwd = 2, 
     mgp = c(2, 1, 0), 
     cex.axis = 1.15)

shadedline(x = z$V1, 
           xaxis = FALSE,
           y0 = z$V2,
           y1 = z$V2 + z$V3, 
           y2 = z$V2 - z$V3,
           title = "",
           line = -1.5,
           adj = 0.02,
           main = paste(title[3]),
           cex.main = 1.25,
           ylab = "CHL resp",
           col = "grey69",
           xlab = "DOY",
           mar = c(5, 3, 1, 3)
           )
axis(side = 1, 
     las = 1, 
     lwd = 2, 
     mgp = c(2, 1, 0), 
     cex.axis = 1.15)
```







