---
title: "Untitled"
author: "Jamie Ash"
date: "2023-02-12"
output: html_document
---

```{r, warning=FALSE}
# Climotology ------------------------------------------------------------------
# Loading packages and all prewriten functions
# setwd("C:\\Users\\james\\Desktop\\jamieslife\\analysis")
source("libraries.R")
source("functions.R")
rasterOptions(maxmemory = 100e+10)
```

Loading the data using ERDDAP
# loading data via opendap had issues with over engineered function
```{r, warning = FALSE, message = FALSE}
# using a two degree box around aloha
lons = c(-175, -125)
lats = c(17, 45)
e = extent(lons, lats)

# Set variables
lat_varid = "lat"
lon_varid = "lon"
sdate = as.Date("2002-01-01")
edate = as.Date("2022-11-01")
var = "CHL"
url = "https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/cmems_obs-oc_glo_bgc-plankton_my_l4-multi-4km_P1M?"
origin = "1900-01-01"

data = nc_open(url, verbose = FALSE, write = FALSE)
lat  = ncvar_get(data, varid = lat_varid)
lon  = ncvar_get(data, varid = lon_varid)
time = ncvar_get(data, varid = "time")
time = as.Date(time, origin = origin)

idx_lat <- which(lat > lats[1] & lat < lats[2])
idx_lon <- which(lon > lons[1] & lon < lons[2])
idx_time <- which(time >= sdate & time <= edate)


idx_ras <- paste("CHL",
                 paste("[", range(idx_time)[1], ":1:", range(idx_time)[2], "]", sep = ""),
                 paste("[", range(idx_lat)[1],  ":1:", range(idx_lat)[2],  "]", sep = ""),
                 paste("[", range(idx_lon)[1],  ":1:", range(idx_lon)[2],  "]", sep = ""),
                 sep = "")

idx_time <- paste("time", paste("[", range(idx_time)[1], ":1:",range(idx_time)[2], "]", sep = ""), sep = "")
idx_lat  <- paste(lat_varid, paste("[", range(idx_lat)[1], ":1:",range(idx_lat)[2],  "]", sep = ""), sep = "")
idx_lon  <- paste(lon_varid, paste("[", range(idx_lon)[1], ":1:",range(idx_lon)[2],  "]", sep = ""), sep = "")
idx <- paste(idx_lat, idx_lon, idx_time, idx_ras, sep = ",")

url <- paste(url, idx, sep = "")

nc_close(data)
rm(data)

data = nc_open(url, verbose = FALSE, write = FALSE)

lat  = ncvar_get(data, varid = lat_varid)
lon  = ncvar_get(data, varid = lon_varid)
time = ncvar_get(data, varid = "time")
time = as.Date(time, origin = origin)
ras  = ncvar_get(data)
nc_close(data)

s = dim(ras)
ras = raster::brick(ras)
ras = t(ras)
ras = setZ(ras, z = as.Date(time, origin = org), name = "time")
extent(ras) = extent(min(lons), max(lons), min(lats), max(lats))

chl = ras
rm(ras)
gc()
```

load chl data
```{r, warning = FALSE}
lons = c(-165, -130)
lats = c(21,   35)
e = extent(lons, lats)

# right now this works because it only has one file in that folder
chl = load_nc(path = "..//data//chl//",
              patt = "chl_glob_l3_20020101_20210105.nc",
              vars = c("lats", "lons", "value", "time"),
              origin = "1970-01-01")
# I saved chl incorrectly
c_e = extent(chl)
extent(chl) = extent(c_e[3], c_e[4], c_e[1], c_e[2])
rm(c_e)
gc()

chl = raster::crop(chl, e)
```

Remove region around Hawaii. CZ for now (EEZ later) 
```{r, eval = TRUE, message = FALSE, warning = FALSE, cache = FALSE}
# chl = bufcoast(chl, 
#                region = "Hawaiian Islands", 
#                path = "../data/USMaritimeLimitsAndBoundariesSHP")
# gc(verbose = FALSE, full = TRUE)
```

Calculate CHL anomaly
```{r, message = FALSE, warning = FALSE, cache = FALSE}
# I receive an warning when detrending for cells of all NA values (hawaii)
# chl anom. calc. needs at least one full year to work correctly
chla = anomalize(chl, detrend = FALSE)
rm(chl)
gc(verbose = FALSE, full = TRUE)
```

correct oreantation
```{r}
chl = subsum(chl)
chla = subsum(chla)
gc()
```



Subset summer and average fro climatology figure
```{r}
cmap   = calc(chla, fun = mean, na.rm = TRUE)
zlim = c(0, 0.012)
cmap = raster::clamp(cmap, zlim[1], zlim[2])
```


Plotting CHL anomaly official plot
```{r, fig.width = 12, fig.height = 7}
# png(file = "../figures/sum_climatology_20221104.png", width = 1200, height = 700)
wdmap = getMap(resolution = "high")
e = extent(cmap)
colmap = cmocean("rain")(50)
lay = matrix(c(1, 2), 
           nrow = 1, 
           byrow = TRUE)

nf = graphics::layout(lay, 
                      widths = c(14, 1))

par(mar = c(4,4,2,1))
image(cmap,
      ylim = c(18, 35),
      xlim = c(-170, -130),
      xlab = "",
      ylab = "",
      zlim = zlim,
      col = colmap,
      axes = FALSE)
plot(wdmap,
     xlim = e[3:4],
     ylim = e[1:2],
     asp = 1,
     bg = "black",
     border = "black",
     col = "black",
     add = TRUE,
     lwd = 4)
axis(side = 2, 
     las = 2, 
     lwd = 2, 
     at = c(18, 22, 26, 30, 34),
     mgp = c(1, 0.75, 0), 
     cex.axis = 1.25)
axis(side = 1, 
     las = 1, 
     lwd = 2,
     mgp = c(2, 1, 0),    
     at = c(-170, -160, -150, -140, -130),
     cex.axis = 1.25)
#title(main = list(main), line = line,  adj = adj, cex.main = cex.main)
title(xlab = "Longitude", line = 2.6, cex.lab = 1.5)
title(ylab = "Latitude", line = 2.6, cex.lab = 1.5)
box(which = "plot",
        lty = "solid",
        lwd = 4,
        col = colvect("grey12", alpha = 0.9))

drawPalette(zlim = zlim,
            col  = colmap,
            plot = TRUE,
            pos  = 4,
            zlab = "",
            cex  = 1.25,
            fullpage = TRUE)

# dev.off()
```

Subset summer and average fro climatology figure
```{r}
# zlim = c(0, 0.2)
# cmap = raster::clamp(temp, zlim[1], zlim[2])

cmap   = calc(chl, fun = mean, na.rm = TRUE)

zlim = c(0.05, 0.10)
cmap = raster::clamp(cmap, zlim[1], zlim[2])
```

Plotting CHL anomaly official plot
```{r, fig.width = 12, fig.height = 7}
#png(file = "../figures/sum_climatology_20221104.png", width = 1200, height = 700)
zlim = c(0.05, 0.10)

wdmap = getMap(resolution = "high")
e = extent(cmap)
colmap = cmocean("rain")(50)
lay = matrix(c(1, 2), 
           nrow = 1, 
           byrow = TRUE)

nf = graphics::layout(lay, 
                      widths = c(14, 1))

par(mar = c(4,4,2,1))
image(cmap,
      ylim = c(18, 35),
      xlim = c(-170, -130),
      xlab = "",
      ylab = "",
      zlim = zlim,
      col = colmap,
      axes = FALSE)
plot(wdmap,
     xlim = c(-170, -130),
     ylim = c(17, 35),
     asp = 1,
     bg = "black",
     border = "black",
     col = "black",
     add = TRUE,
     lwd = 4)
axis(side = 2, 
     las = 2, 
     lwd = 2, 
     at = c(18, 22, 26, 30, 34),
     mgp = c(1, 0.75, 0), 
     cex.axis = 1.25)
axis(side = 1, 
     las = 1, 
     lwd = 2,
     mgp = c(2, 1, 0),    
     at = c(-170, -160, -150, -140, -130),
     cex.axis = 1.25)
#title(main = list(main), line = line,  adj = adj, cex.main = cex.main)
title(xlab = "Longitude", line = 2.6, cex.lab = 1.5)
title(ylab = "Latitude", line = 2.6, cex.lab = 1.5)
box(which = "plot",
        lty = "solid",
        lwd = 4,
        col = colvect("grey12", alpha = 0.9))

drawPalette(zlim = c(0.03, 0.12),
            col  = colmap,
            plot = TRUE,
            pos  = 4,
            zlab = "",
            cex  = 1.25,
            fullpage = TRUE)
#dev.off()
```








---------------------------------------
unofficial plots


```{r}
sdate = as.Date("2018-06-01")
edate = as.Date("2018-11-15")
chl_2018 = timesnip(chl,  sdate, edate)
chla_2018 = timesnip(chla,  sdate, edate)
chla_2018 = bufcoast(chla_2018)
```

For CHL anomaly as well?
```{r}
zlim = c(0.02, 0.15)
cmap_2018   = calc(chl_2018, fun = mean, na.rm = TRUE)
cmap_2018 = raster::clamp(cmap_2018, lower = zlim[1], upper=zlim[2])

zlim = c(-0.005, 0.025)
cmap_2018_anom   = calc(chla_2018, fun = mean, na.rm = TRUE)
cmap_2018_anom = raster::clamp(cmap_2018_anom, lower=zlim[1], upper=zlim[2])

```

Plotting CHL 2018 official plot
```{r, fig.width = 12, fig.height = 7}
#png(file = "../figures/sum_climatology_20221104.png", width = 1200, height = 700)
wdmap = getMap(resolution = "high")
e = extent(cmap_2018)
zlim = c(0.03, 0.15)

colmap = cmocean("rain")(50)
lay = matrix(c(1, 2), 
           nrow = 1, 
           byrow = TRUE)

nf = graphics::layout(lay, 
                      widths = c(14, 1))

par(mar = c(4,4,2,1))
image(cmap_2018,
      ylim = c(18, 35),
      xlim = c(-170, -130),
      xlab = "",
      ylab = "",
      zlim = zlim,
      col = colmap,
      axes = FALSE)
plot(wdmap,
     xlim = c(-170, -130),
     ylim = c(17, 35),
     asp = 1,
     bg = "black",
     border = "black",
     col = "black",
     add = TRUE,
     lwd = 4)
axis(side = 2, 
     las = 2, 
     lwd = 2, 
     at = c(18, 22, 26, 30, 34),
     mgp = c(1, 0.75, 0), 
     cex.axis = 1.25)
axis(side = 1, 
     las = 1, 
     lwd = 2,
     mgp = c(2, 1, 0),    
     at = c(-170, -160, -150, -140, -130),
     cex.axis = 1.25)
#title(main = list(main), line = line,  adj = adj, cex.main = cex.main)
title(xlab = "Longitude", line = 2.6, cex.lab = 1.5)
title(ylab = "Latitude", line = 2.6, cex.lab = 1.5)
box(which = "plot",
        lty = "solid",
        lwd = 4,
        col = colvect("grey12", alpha = 0.9))

drawPalette(zlim = zlim,
            col  = colmap,
            plot = TRUE,
            pos  = 4,
            zlab = "",
            cex  = 1.25,
            fullpage = TRUE)
#dev.off()
```

Plotting CHL 2018 official plot
```{r, fig.width = 12, fig.height = 7}
#png(file = "../figures/sum_climatology_20221104.png", width = 1200, height = 700)
wdmap = getMap(resolution = "high")
e = extent(cmap_2018_anom)
zlim = c(-0.005, 0.025)

colmap = cmocean("rain")(50)
lay = matrix(c(1, 2), 
           nrow = 1, 
           byrow = TRUE)

nf = graphics::layout(lay, 
                      widths = c(14, 1))

par(mar = c(4,4,2,1))
image(cmap_2018_anom,
      ylim = c(18, 37),
      xlim = c(-170, -130),
      xlab = "",
      ylab = "",
      zlim = zlim,
      col = colmap,
      axes = FALSE)
plot(wdmap,
     xlim = c(-170, -130),
     ylim = c(17, 35),
     asp = 1,
     bg = "black",
     border = "black",
     col = "black",
     add = TRUE,
     lwd = 4)
axis(side = 2, 
     las = 2, 
     lwd = 2, 
     at = c(18, 22, 26, 30, 34),
     mgp = c(1, 0.75, 0), 
     cex.axis = 1.25)
axis(side = 1, 
     las = 1, 
     lwd = 2,
     mgp = c(2, 1, 0),    
     at = c(-170, -160, -150, -140, -130),
     cex.axis = 1.25)
#title(main = list(main), line = line,  adj = adj, cex.main = cex.main)
title(xlab = "Longitude", line = 2.6, cex.lab = 1.5)
title(ylab = "Latitude", line = 2.6, cex.lab = 1.5)
box(which = "plot",
        lty = "solid",
        lwd = 4,
        col = colvect("grey12", alpha = 0.9))

drawPalette(zlim = zlim,
            col  = colmap,
            plot = TRUE,
            pos  = 4,
            zlab = "",
            cex  = 1.25,
            fullpage = TRUE)
#dev.off()
```

